\blue{Numerical Investigation of Multi-Fluid Flows through Dual Saturated-Unsaturated Porous Media - change the title to something close to the main topic of the manuscript ... something with viscous flow instability or viscous fingering)}

### 2017-03-03 Papers/ViscousInstabilities/Manuscript_ViscousInstabilities_KC.tex
final 1

\begin{abstract} 
\KCnote{ \blue{ Multiphase flows in porous media is an interesting and important subject in fields ranging from oil recovery to medicine. From physics standpoint, the complex structures observed in multiphase flows, also know as fingers, are especially interesting. Much work both experimental and numerical has been done on invasion systems where one fluid is injected into a system filled with another fluid. These patterns are causing an instability known as Safmman-Taylor or Viscous fingering. Continuity equation and the Darcy's law govern the processes, while pressure, velocity, saturation and permeability are quantities that define the behaviour of the flow. A novel control finite element method (CVFEM) is used to discretize the governing equations that will solve the multi-fluid porous media flow model. The current model is the basis of a general purpose flow simulator that is equipped with anisotropic mesh adaptivity techniques to increase the computational accuracy of numerical methods used in geophysical fluid mechanics and tackle the difficulties that commercial resevoir simulators may rise when using low order finite volume/differences methods based on structured grids. The governing multiphase porous media flow equations are solved using an adaptive and unstructured mesh. The model is based upon two key numerical characteristics: (a) the family of $P_{n}DG-P_{n+1}$ finite element pairs and (b) a consistent overlapping control volume finite element method (CVFEM) formulation.
%Moreover, the mesh adaptivity algorithm employs multi-constraints on element size in different regions of the porous medium to resolve multi-scale transport phenomena. The scheme captures the key characteristics of the flow while its applicable to heterogeneous porous media and geological formations. We conclude that understanding the multiphase flow properties will allow us to efficiently facilitate more efficient oil \& gas recovery, efficient contaminant removal and better planning of carbon capture and storage (CCS) projects.
In this paper the use of dynamic mesh adaptivity to capture the viscous fingering in reservoir flows is introduced, as one of the methods that will allow the user to focus the effort where is needed, minimizing both the CPU time and the numerical diffusion. Both the evolution of visualisation techniques through the extensive use of parallel computing as well as the advances in modelling and simulation of porous media flows may help to address the importance of this phenomenon which can be found in petroleum, chemical and environmental operations.
The aim of this paper is to extend and further apply the formulation (embedded in the Fluidity software framework) to multiphase flows in heterogeneous porous media combined with either uniform or random generated permeability distribution...}}

@@

\medskip
Figure \ref{fig:3testcase_a} illustrates another set of simulations for when another region (inlet) of fluid $1$ ins introduced into the formation. This time fluid $1$ is injected from both the left hand side and the upper left corner of the domain, at a velocity u=$0.5$ while the set up of the problem remains the same, the mobility ratio is $10$, and the adaptive and ustructured mesh is implied. The top layer is saturated with phase $2$, while the rest of the domain has a saturation of $0.1$. The major difference is located in the velocity fields (see Fig.\ref{fig:vel_magn2}) as the presence of this extra inlet increases the flow at the top layers. At this point is proper to mentioned the extreme case, see Fig.\ref{fig:4testcase}, when mobility ratio is even an order of magnitude higher, when MR=$100$. Even though a this case is unlikely to happen as the mobility ration of the existing fluids in the industry, it is a test-case that can illustrate the higher, faster and more profound development of instabilities. Also by focusing the grid resolution in the saturation front, \citet{christou2015}  demonstrated the explicit formation of fingers formed due to the instability in the pressure ﬁeld and triggered by viscosity ratio. A comparison has also been made by using a low resolution element, the $P_{1}DG-P_{1}DG$. No fingers could be observed due to mesh coarseness (fingers were observed in ﬁner uniform and adaptive meshes).

@@

\begin{comment}
\begin{figure}
\centering
\subfigure[fixed mesh, t=500]{\includegraphics[width=0.75\textwidth]{./Pics1/5reg_dinlet_fixed_500.pdf}}\\[2mm]%
\subfigure[adaptive mesh, t=500]{\includegraphics[width=0.75\textwidth]{./Pics1/5reg_dinlet_adapt_500.pdf}}%
\caption{for t=$500$ there is a comparison between fixed mesh(a) and adaptive mesh(b).}
\label{fig:3testcase_b}
\end{figure}

@@

\begin{figure}
\centering
\subfigure[fixed mesh, t=end]{\includegraphics[width=0.75\textwidth]{./Pics1/5reg_dinlet_fixed_end.pdf}}\\[2mm]%
\subfigure[adaptive mesh, t=end]{\includegraphics[width=0.75\textwidth]{./Pics1/5reg_dinlet_adapt_end.pdf}}%
\caption{for t=end of simulation, there is a comparison between fixed mesh(a) and adaptive mesh(b).}
\label{fig:3testcase_d}
\end{figure}
\end{comment}

>>>>

\begin{abstract} 
A set of simulations is presented so as to model the viscous fingering-flow instability using a novel method for modeling multi-phase flows in porous media. Starting from the Hele-Shaw experimental which is mathematically analogous to the two-dimensional flow in porous medium, test-cases of both homogenous and heterogeneous computational domains, under different conditions are listed. The development of finger structures is explained by taking into account different maps of permeabilities, mobility ratios and surface tension. A novel control finite element method (CVFEM) is used to discretize the governing equations that will solve the multi-fluid porous media flow model. The governing multiphase porous media flow equations are solved using an adaptive and unstructured mesh. In this paper the use of dynamic mesh adaptivity to capture the viscous fingering in reservoir flows is introduced, as one of the methods that will allow the user to focus the effort where is needed, minimizing both the CPU time and the numerical diffusion. It was found that the initial state is the decisive factor in the development of the process up to large times 
% Multiphase flows in porous media is an interesting and important subject in fields ranging from oil recovery to medicine. From physics standpoint, the complex structures observed in multiphase flows, also know as fingers, are especially interesting. Much work both experimental and numerical has been done on invasion systems where one fluid is injected into a system filled with another fluid. These patterns are causing an instability known as Safmman-Taylor or Viscous fingering. Continuity equation and the Darcy's law govern the processes, while pressure, velocity, saturation and permeability are quantities that define the behaviour of the flow.  The current model is the basis of a general purpose flow simulator that is equipped with anisotropic mesh adaptivity techniques to increase the computational accuracy of numerical methods used in geophysical fluid mechanics and tackle the difficulties that commercial resevoir simulators may rise when using low order finite volume/differences methods based on structured grids. The governing multiphase porous media flow equations are solved using an adaptive and unstructured mesh. The model is based upon two key numerical characteristics: (a) the family of $P_{n}DG-P_{n+1}$ finite element pairs and (b) a consistent overlapping control volume finite element method (CVFEM) formulation.
%Moreover, the mesh adaptivity algorithm employs multi-constraints on element size in different regions of the porous medium to resolve multi-scale transport phenomena. The scheme captures the key characteristics of the flow while its applicable to heterogeneous porous media and geological formations. We conclude that understanding the multiphase flow properties will allow us to efficiently facilitate more efficient oil \& gas recovery, efficient contaminant removal and better planning of carbon capture and storage (CCS) projects.
%In this paper the use of dynamic mesh adaptivity to capture the viscous fingering in reservoir flows is introduced, as one of the methods that will allow the user to focus the effort where is needed, minimizing both the CPU time and the numerical diffusion. Both the evolution of visualisation techniques through the extensive use of parallel computing as well as the advances in modelling and simulation of porous media flows may help to address the importance of this phenomenon which can be found in petroleum, chemical and environmental operations.
The aim of this paper is to extend and further apply the formulation (embedded in the Fluidity software framework) to multiphase flows in heterogeneous porous media combined with either uniform or random generated permeability distribution.

@@

\medskip
Figure \ref{fig:3testcase_a} illustrates another set of simulations for when another region (inlet) of fluid $1$ ins introduced into the formation. This time fluid $1$ is injected from both the left hand side and the upper left corner of the domain, at a velocity u=$0.5$ while the set up of the problem remains the same, the mobility ratio is $10$, and the adaptive and ustructured mesh is implied. The top layer is saturated with phase $2$, while the rest of the domain has a saturation of $0.1$. The major difference is located in the velocity fields (see Fig.\ref{fig:vel_magn2}) as the presence of this extra inlet increases the flow at the top layers. At this point is proper to mentioned the extreme case, see Fig.\ref{fig:4testcase}, when mobility ratio is even an order of magnitude higher, when MR=$100$. Even though a this case is unlikely to happen as the mobility ration of the existing fluids in the industry, it is a test-case that can illustrate the higher, faster and more profound development of instabilities. Also by focusing the grid resolution in the saturation front, \citet{christou_2015}  demonstrated the explicit formation of fingers formed due to the instability in the pressure ﬁeld and triggered by viscosity ratio. A comparison has also been made by using a low resolution element, the $P_{1}DG-P_{1}DG$. No fingers could be observed due to mesh coarseness (fingers were observed in ﬁner uniform and adaptive meshes).

@@

%\begin{comment}
\begin{figure}
\centering
\subfigure[fixed mesh, t=500]{\includegraphics[width=0.75\textwidth]{./Pics1/5reg_dinlet_fixed_500.pdf}}\\[2mm]%
\subfigure[adaptive mesh, t=500]{\includegraphics[width=0.75\textwidth]{./Pics1/5reg_dinlet_adapt_500.pdf}}%
\caption{for t=$500$ there is a comparison between fixed mesh(a) and adaptive mesh(b).}
\label{fig:3testcase_b}
\end{figure}

@@

\begin{figure}
\centering
\subfigure[fixed mesh, t=end]{\includegraphics[width=0.75\textwidth]{./Pics1/5reg_dinlet_fixed_end.pdf}}\\[2mm]%
\subfigure[adaptive mesh, t=end]{\includegraphics[width=0.75\textwidth]{./Pics1/5reg_dinlet_adapt_end.pdf}}%
\caption{for t=end of simulation, there is a comparison between fixed mesh(a) and adaptive mesh(b).}
\label{fig:3testcase_d}
\end{figure}
%\end{comment}
